---
title: "CU-specific performance metrics"
author: "Cam Freshwater"
date: "December 2, 2018"
output: html_document
---

In compCVOpModels.Rmd I noticed that patterns in CU-specific PMs varied among CUs. Here I'm focusing on outputs from the reference op model to explore how CUs differ in effects of greater CVc (under median synch) and greater synchrony (under median sigma). Note that while the interaction between synch and CVc (as well as various operating models) will also vary among CUs, it is going to be difficult to fully explore this and incorporate the findings into the synchrony manuscript. Therefore the goal here is simply to get an understanding of variability among CUs and potential mechanisms.

```{r readLibraries, message=FALSE, warning=FALSE, include=FALSE}
listOfPackages <- c("plyr", "here", "parallel", "doParallel", "foreach", 
                    "reshape2", "tidyverse", "gsl", "tictoc", "stringr", 
                    "synchrony", "zoo", "Rcpp", "RcppArmadillo", "sn", 
                    "sensitivity", "mvtnorm", "forcats", "ggpubr", "viridis", 
                    "samSim")

here <- here::here

newPackages <- listOfPackages[!(listOfPackages %in% 
                                  installed.packages()[ , "Package"])]
if(length(newPackages)) install.packages(newPackages)
lapply(listOfPackages, require, character.only = TRUE)
```


```{r importData, message=FALSE, warning=FALSE, include=FALSE}
simPar <- read.csv(here("data/sox/fraserOMInputs_varyCorr.csv"),
                   stringsAsFactors = F)
cuPar <- read.csv(here("data/sox/fraserCUpars.csv"), stringsAsFactors = F)
srDat <- read.csv(here("data/sox/fraserRecDatTrim.csv"), stringsAsFactors = F)
catchDat <- read.csv(here("data/sox/fraserCatchDatTrim.csv"), 
                     stringsAsFactors = F)
ricPars <- read.csv(here("data/sox/rickerMCMCPars.csv"), stringsAsFactors = F)
larkPars <- read.csv(here("data/sox/larkinMCMCPars.csv"), stringsAsFactors = F)
tamFRP <- read.csv(here("data/sox/tamRefPts.csv"), stringsAsFactors = F)

simParTrim <- subset(simPar, 
                     scenario == "lowSig" | scenario == "medSig" | 
                       scenario == "highSig"
                     )
scenNames <- unique(simParTrim$scenario)
dirNames <- sapply(scenNames, function(x) paste(x, unique(simParTrim$species), 
                                                sep = "_"))
```


```{r runSimulations, eval=FALSE, include=FALSE}
## Code chunk coded out because generally these simulation runs will already be present.

## Define simulations to be run
# nTrials <- 75
# 
# for (i in seq_along(dirNames)) {
#   dirName <- dirNames[i]
#   d <- subset(simParTrim, scenario == scenNames[i])
#   simsToRun <- split(d, seq(nrow(d)))
#   Ncores <- detectCores()
#   cl <- makeCluster(Ncores - 3) #save two cores
#   registerDoParallel(cl)
#   clusterEvalQ(cl, c(library(MASS),
#                   library(here),
#                   library(sensitivity),
#                   library(mvtnorm),
#                   library(scales), #shaded colors for figs
#                   library(viridis), #color blind gradient palette
#                   library(gsl), 
#                   library(dplyr),
#                   library(Rcpp),
#                   library(RcppArmadillo),
#                   library(sn),
#                   library(samSim)))
#   #export custom function and objects
#   clusterExport(cl, c("simsToRun", "recoverySim", "cuPar", "dirName", "nTrials",
#                       "catchDat", "srDat", "ricPars", "dirName", "larkPars", 
#                       "tamFRP"), envir = environment()) 
#   tic("run in parallel")
#   parLapply(cl, simsToRun, function(x) {
#     recoverySim(x, cuPar, catchDat = catchDat, srDat = srDat, 
#                 variableCU = FALSE, ricPars, larkPars = larkPars, 
#                 tamFRP = tamFRP, cuCustomCorrMat = NULL, dirName = dirName, 
#                 nTrials = nTrials, multipleMPs = FALSE)
#   })
#   stopCluster(cl) #end cluster
#   toc()
# }
```


```{r cuSpecificPMs, echo=TRUE}
omNames <- rep(c("ref"), each = 3)
sigNames <- rep(c("low", "med", "high"), length.out = length(omNames))
stkNames <- genOutputList(dirNames[1], 
                          agg = FALSE)[["medSynch_TAM"]][["stkName"]]
nCUs <- length(stkNames)

#First generate data and plots for different sigma treatments
plotDat <- NULL
for (i in seq_along(dirNames)) { #make dataframe
  cuList <- genOutputList(dirNames[i], agg = FALSE)[["medSynch_TAM"]]
  nTrials <- nrow(cuList[["medSpawners"]])
  spwnDat <- data.frame(om = rep(omNames[i], length.out = nTrials * nCUs),
                        sigma = rep(sigNames[i], length.out = nTrials * nCUs)
  )
  spwn <- cuList[["medSpawners"]] %>%
    as.data.frame() 
  colnames(spwn) <- stkNames
  spwnLong <- spwn %>% 
    gather(key = cu, value = spawners) %>% 
    mutate(cu = as.factor(cu))
  plotDat <- rbind(plotDat, cbind(spwnDat, spwnLong))
}

axisSize = 15; dotSize = 3.5; lineSize = 0.8
q <- ggplot(plotDat, aes(x = sigma, y = spawners, alpha = sigma)) +
  geom_violin(draw_quantiles = c(0.5), position = position_dodge(width = 0.75),
              fill = "grey30") +
  scale_alpha_manual(name = "Sigma OM", values = c(1, 0.65, 0.3),
                     labels = c("low" = expression(paste("0.75", sigma)),
                                "med" = expression(paste("1.0", sigma)),
                                "high" = expression(paste("1.25", sigma)))) +
  guides(alpha = guide_legend(override.aes = list(fill = "grey30"))) +
  labs(y = "Median Spawner\n Abundance (millions)",
       x = "Sigma OM") +
  theme_sleekX(facetSize = 1.2, axisSize = 12, legendSize = 0.85) +
  facet_wrap(~cu, scales = "free_y")
q
```

The majority of CUs exhibit declines in abundance with increasing variability, but the strength of this effect varies. Notable exceptions are Gates/Harrison/Nadina/Seymour (medians stable even though variance increases) and Raft/Cultus (median increases along with variance).

Repeat previous analysis with trends in synchrony instead (should be much more consistent).

```{r cuSpecificPMsSynch, echo = TRUE}
omNames <- rep(c("ref"), each = 3)
#Make sure order of synchrony names is correct based on contents of cuList
synchNames <- rep(c("high", "low", "med"), length.out = length(omNames))
stkNames <- genOutputList(dirNames[1], 
                          agg = FALSE)[["medSynch_TAM"]][["stkName"]]
nCUs <- length(stkNames)

#First generate data and plots for different sigma treatments
plotDat <- NULL
for (i in seq_along(synchNames)) { #make dataframe
  #Focus on median sigma treatment (i.e. dirNames2)
  cuList <- genOutputList(dirNames[2], agg = FALSE)[[i]]
  nTrials <- nrow(cuList[["medSpawners"]])
  spwnDat <- data.frame(om = rep(omNames[i], length.out = nTrials * nCUs),
                        synch = rep(synchNames[i], length.out = nTrials * nCUs)
  )
  spwn <- cuList[["medSpawners"]] %>%
    as.data.frame() 
  colnames(spwn) <- stkNames
  spwnLong <- spwn %>% 
    gather(key = cu, value = spawners) %>% 
    mutate(cu = as.factor(cu))
  plotDat <- rbind(plotDat, cbind(spwnDat, spwnLong))
}

q <- ggplot(plotDat, aes(x = synch, y = spawners, alpha = synch)) +
  geom_violin(draw_quantiles = c(0.5), position = position_dodge(width = 0.75),
              fill = "grey30") +
  scale_alpha_manual(name = "Synch OM", values = c(1, 0.65, 0.3)) +
  guides(alpha = guide_legend(override.aes = list(fill = "grey30"))) +
  labs(y = "Median Spawner\n Abundance (millions)",
       x = "Synch OM") +
  theme_sleekX(facetSize = 1.2, axisSize = 12, legendSize = 0.85) +
  facet_wrap(~cu, scales = "free_y")
q
```

As expected effects of synchrony are basically the same across CUs (note that this is only expected for the reference case; alternative OMs explored in cuSpecificPMs_synchAltOMs.Rmd).

How are these related to alpha and beta values?

```{r plotBeta, echo=TRUE}
parDat <- data.frame(cu = cuList$stkName,
                      medBeta = cuList$medBeta[1,],
                      medAlpha = cuList$medAlpha[1,],
                     sigma = cuPar$sigma
                     ) %>% 
  melt() %>% 
  dplyr::rename("par" = "variable") %>% 
  mutate(decline = case_when(
           cu %in% c("Clts", "Gats", "Hrrs", "Nadn", "Raft", "Symr") ~ "no",
           TRUE ~ "yes"
         ),
         model = case_when(
           cu %in% c("E.St", "L.St", "Qsnl", "Symr", "L.Sh") ~ "larkin",
           TRUE ~ "ricker"
         ))

colPal <- c("#b2182b", "#2166ac")
ggplot(parDat, aes(x = cu, y = value, colour = decline, shape = model)) +
  geom_point() +
  scale_color_manual(name = "Decline", values = colPal) +
  scale_shape_discrete(name = "SR Model") +
  theme_sleekX() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~par, scales = "free_y")

```

Only patterns that I observe are that stocks that do not exhibit a decline with greater variability (red) typically have lower alpha values. Thus it seems as though the pattern must be driven by some interaction between multiple parameters? Again not sure if we've fully identified the drivers, but I'm unsure of where to go from here.